Ignition of Fuel Beds by Firebrands

### 2022-04-11 litreview.tex
Updated species study and literature review
Completed the section on hot spot theory

    In order to identify the controlling parameters of ignition and create a model of quantitative accuracy the processes that occur leading up to ignition must be identified. Figure~\ref{fig:ignitionDiagram} illustrates the different factors that can influence the ignition process. To better understand how these parameters effect the ignition process can be considered as a series of three processes.  The first is the heat transfer to the fuel bed from the ember. The second is the pyrolysis of the fuel bed material and the subsequent release of pyrolyzates into the air above the fuel bed. The third step is the mixing of the pyrolyzates with air above the fuel bed and subsequent gas phase ignition. It is important to note that all three of these processes may occur simultaneously throughout the ignition process and are often interconnected. For example, as heat is transferred from the firebrand to the fuel bed the thermal conductivity of the fuel bed changes with temperature both before and after pyrolysis begins~\cite{Fjellerup2003}. Furthermore, the contact area between the firebrand and fuel continuously changes as mass is released due to pyrolysis. Due to the interconnection of the processes that occur during ignition and the inherent variability of fuel beds and biomass fuels in general quantification of parameters that influence ignition is quite challenging. Despite these challenges results of previous research have identified influential parameters that have consistently been observed to control ignition. This work considers these parameters as belonging to one of three categories namely, thermo-physical properties of the fuel bed, energy supplied by the firebrand, and environmental conditions near the ignition point. Detailed reviews of each category are provided in the relevant manuscripts and are not reproduced here. The remaining sections of this review address the predictive models of ignition that have been created from the current state of knowledge. The accuracy and appilicability of each model is discussed in turn and the remaining challenges are identified as they have motivated the objectives of this work.

         \begin{figure}[hpbt]
            \centering
                \includegraphics[width=0.75\textwidth]{Figures/dissertationBed.pdf}
            \caption{Diagram illustrating potential parameters that may influence ignition of a fuel bed.}
            \label{fig:ignitionDiagram}
        \end{figure}
    A significant number of ignition criteria and models have been proposed to predict fuel bed ignition, however no model has provided a general quantitatively accurate prediction of ignition. 
    \section{Current Ignition Models}
        \begin{itemize}
            \item Hot spot theory
            \item Critical mass flux
            \item Energy content of the particle
            \item Numerical methods
            \item Analytic correlations
        \end{itemize}

>>>>

    In order to identify the controlling parameters of ignition and create a model of quantitative accuracy the processes that occur leading up to ignition must be identified. Figure~\ref{fig:ignitionDiagram} illustrates the different factors that can influence the ignition process. To better understand how these parameters effect the ignition process can be considered as a series of three processes.  The first is the heat transfer to the fuel bed from the ember. The second is the pyrolysis of the fuel bed material and the subsequent release of pyrolyzates into the air above the fuel bed. The third step is the mixing of the pyrolyzates with air above the fuel bed and subsequent gas phase ignition. It is important to note that all three of these processes may occur simultaneously throughout the ignition process and are often interconnected. For example, as heat is transferred from the firebrand to the fuel bed the thermal conductivity of the fuel bed changes with temperature both before and after pyrolysis begins~\cite{Fjellerup2003}. Furthermore, the contact area between the firebrand and fuel continuously changes as mass is released due to pyrolysis. Due to the interconnection of the processes that occur during ignition and the inherent variability of fuel beds and biomass fuels in general quantification of parameters that influence ignition is quite challenging. Despite these challenges results of previous research have identified influential parameters that have consistently been observed to control ignition. This work considers these parameters as belonging to one of three categories namely, thermo-physical properties of the fuel bed, energy supplied by the firebrand, and environmental conditions near the ignition point. Detailed reviews of each category are provided in the relevant manuscripts and are not reproduced here. The remaining sections of this review address the predictive models of ignition that have been created from the current state of knowledge. The accuracy and appilicability of each model is discussed in turn and the remaining challenges are identified as they have motivated the objectives of this work. 
        \begin{figure}[hpbt]
            \centering
                \includegraphics[width=0.75\textwidth]{Figures/dissertationBed.pdf}
            \caption{Diagram illustrating potential parameters that may influence ignition of a fuel bed.}
            \label{fig:ignitionDiagram}
        \end{figure}

    Perhaps the first model to be considered for the ignition of wildland fuels is the hot spot theory. The hot spot model considers the ignition of an explosive (i.e., fuel bed) by a heat source (firebrand) by determining the energy required for a thermally explosive reaction in the explosive~\cite{Zinn1962InitiationSpots, Thomas1965}. The minimum hot particle size for ignition of a fuel bed, as presented by Hadden et al.~\cite{Hadden2011}, is shown in Equation~\ref{eqn:hotspot}.

        \begin{equation}
            r_{cr} = \delta_{cr}\sqrt{\frac{k}{\rho A \Delta H}\frac{RT^{2}_{p0}}{E} exp\left(\frac{E}{RT_{p0}} \right)}
            \label{eqn:hotspot}
        \end{equation}
    In this equation $\delta_{cr}$ is defined as the critical Frank-Kamenetskii hot spot parameter as determined from calculations~\cite{Goldshlager} and is dependent on the ratio of volumetric heat capacities of the ignition source and the material, the activation temperature of the material, and the temperature of the hot particle. Ultimately, Equation~\ref{eqn:hotspot} defines the particle energy needed to raise the temperature of a fuel bed region above the activation temperature to enable thermal runaway and subsequent ignition. While a relatively simple, the hot spot theory has some limitations that prevent quantitative predictions of ignition in a wildfire setting. A discussion of the shortcomings of the hot spot theory are presented by Thomas~\cite{Thomas1965} and more recently presented by Hadden et al.~\cite{Hadden2011} with respect to wildfires. The two most significant shortcomings addressed by the previous works are neglecting heat losses to the surroundings and considering the pyrolysis and ignition phenomena as single step infinitely fast reactions. Hadden et al. attributes these shortcomings as a significant source of error for quantitative ignition predictions. This assertion is reinforced in a review by Fernandez-Pello who concluded that in instances where hot spot theory has been applied only qualitatively accurate predictions of ignition have been obtained. Ultimately, the hot spot theory does not capture the complexity and variability inherent to ignition.

    A significant number of ignition criteria and models have been proposed to predict fuel bed ignition, however no model has provided a general quantitatively accurate prediction of ignition. 
    \section{Current Ignition Models}
        \begin{itemize}
            \item Hot spot theory (Hadden 2011)
            \item Critical mass flux (Nelson 1995)
            \item Energy content of the particle
            \item Numerical methods
            \item Analytic correlations
        \end{itemize}
